---
title: 'Juvenile snowshoe hares in Bonanza Creek: exploratory report'
author: "Anna Zauner"
date: "11/4/2021"
output: 
  html_document: 
    theme: cerulean
    toc: yes
---
# Introduction 

This report provides an mini exploration of differences in juvenile snowshoe hare trap counts and weights from 1999-2012 in Bonanza Creek Experimental Forest in Interior Alaska (Kielland et al. 2017). This capture-recapture study explores the snowshoe hare, which is a "keystone" prey species in the northern boreal forests in 3 different sites (Black Spruce, Mature and Riparian) (Kielland et al. 2017). This report first explores the juvenile snowshoe hare trap counts annually from 1998 to 2012. Then explores the juvenile snowshoe hare weights by sex and site. Followed by a more focused comparison of weights between male and female juvenile hares. Lastly, explores the relationship between juvenile weights and hind foot length. 

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Set up 
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
library(effsize)
library(broom)
library(ggbeeswarm)
```

```{r}
# Read in the data
bonanza_hares_data <- read_csv(here("data","bonanza_hares.csv")) %>% 
  clean_names() %>% 
  mutate(date = mdy(date))
```

# Data and anyalses

The data on snowshoe hares was collected by capture-recapture by Dr. Knut Kielland and colleagues at the Bonanza Creek Experimental Forest in Fairbanks, Alaska (Kielland et al. 2017). The data contain 3,197 observations of snowshoe hares, collected from 1999-2012 over three sites (Black Spruce, Mature and Riparian). Following exploratory visualization, juvenile weights (grams) and hind foot lengths (millimeters) are compared by two-sample t-test using a significance level ($\alpha$) of 0.05 and differences between groups are described by Cohen’s *d* effect size. The relationship between weight and hind foot length is explored by simple linear regression (All analyses are in R version 4.0.2 using RStudio version 1.4.1717)
 
# Juvenile hare initial investigation 

## Section 1: Annual juvenile hare trap counts

```{r fig.align='center'}

# Wrangle the data to include juvenile hares, group by year, and summarize

annual_juvenile_trap_counts <- bonanza_hares_data %>%
  filter(age == "j") %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(trap_counts = n())

# Visualize the data with geom_bar 

ggplot(data = annual_juvenile_trap_counts, aes(x = year, y = trap_counts))+
  geom_bar(stat = "identity", fill = "blue")+
  labs(x = "Year",
       y = "Juvenile hare trap counts",
       title = "Annual juvenile hare trap counts (1999-2012) (ZAUNER)")+
  theme_classic()+
  theme(plot.caption = element_text(hjust = 0.5, size = 8, face = "bold"),
        plot.title = element_text(hjust = 0.5, face = "bold"))+
  scale_x_continuous(breaks = c(1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012))


```
**Figure 1:** Counts of juvenile hare trappings during each year of the study from 1998 to 2012 in the Bonanza Creek Experimental Forest. The trap counts peak in 1999 at 126 hares data from Kielland et al.(2017).

```{r}

# Creat stored summary statistics table 

annual_counts_summary <- annual_juvenile_trap_counts %>% 
  summarize(mean = mean(trap_counts),
            median = median(trap_counts),
            minimum = min(trap_counts),
            maximum = max(trap_counts))


```

### Figure 1 takeaways

There is a general trend of decreasing counts of juvenile hare trappings from 1999-2012, there is a large spread of annual juvenile hare trap counts, ranging from `r annual_counts_summary$minimum` to `r annual_counts_summary$maximum`. The mean count of juvenile hares trapped annually is `r annual_counts_summary$mean` and median is `r annual_counts_summary$median`. The changes from year to year show no consistent trend, other than a large drop after the maximum count year in 1999, this could be due to the fact that these are absolute counts and are impacted by the effort of researchers. I would suggest standardizing the juvenile hare population by randomly selecting uniform sample sizes for each year. 

## Section 2: Visualize juvenile hare weights

```{r fig.align='center'}
# Wrangle the data 
juvenile_weights_sex_site <- bonanza_hares_data %>% 
  filter(age == "j") %>% 
  select(sex, weight, grid) %>% 
  mutate(sex = case_when(
    sex == "m" ~ "Male",
    sex == "f" ~ "Female",
    is.na(sex) == TRUE ~ "Unknown")) %>% 
  mutate(grid_full = case_when(
    grid == "bonbs" ~ "Black Spruce",
    grid == "bonmat" ~ "Mature",
    grid == "bonrip" ~ "Riparian"
  ))

# Visualize the data using a beeswarm graph 
ggplot(data = juvenile_weights_sex_site, aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex, show.legend = "None"), cex = 3) +
  scale_color_manual(values = c("goldenrod2","cyan4","coral")) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  theme_minimal() +
  labs(x = "Sex",
       y = "Weight (g)",
       title = "Juvenile hare weight (g) by sex and site (ZAUNER)")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title = element_text(face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 0.5))+
  facet_wrap(~grid_full)
```
**Figure 2:** Juvenile hare weight observations (g) by sex (male, female and unknown) and site (Black Spruce, Mature and Riparian). Gold (female), teal (male) and coral (unknown) points indicate observations for weight (grams) of a juvenile hare. Black point indicates the mean value for each sex. Data: Kielland et al. (2017).

```{r}
# Under the hood statistics table to reference
juvenile_weights_sex_site_ref <- juvenile_weights_sex_site %>% 
  group_by(grid_full, sex) %>% 
  summarize(sample_size = n(),
            meadian = median(weight, na.rm = TRUE),
            mean = mean(weight, na.rm = TRUE),
            min = min(weight, na.rm = TRUE),
            max = max(weight, na.rm = TRUE),
            sd = sd(weight, na.rm = TRUE))

```

```{r}
# Summary statistics

juvenile_weights_sex_site_summary <- juvenile_weights_sex_site %>% 
  group_by(grid_full, sex) %>% 
  summarize(sample_size = n(),
            meadian = median(weight, na.rm = TRUE),
            mean = mean(weight, na.rm = TRUE, digits = 1),
            min = min(weight, na.rm = TRUE),
            max = max(weight, na.rm = TRUE),
            sd = sd(weight, na.rm = TRUE)) %>% 
  kbl(caption = "Table 1: Descriptive statistics (sample size, median, mean, minimum, maximum and standard deviation) for male, female and unknown juvenile hare weights (g). Data: Kielland et al. (2017)",
      col.names = c("Site",
                    "Sex",
                    "Sample size",
                    "Median (g)",
                    "Mean (g)",
                    "Min (g)",
                    "Max (g)",
                    "Standard deviation (g)")) %>% 
  kable_minimal(html_font = "Georgia", font_size = 12, full_width = FALSE) %>% 
  kable_material(c("striped", "hover", "condensed"))
  
juvenile_weights_sex_site_summary

```

### Juvenile hare weights and descriptive statistics

There is a visible difference in sample sizes of juvenile hares between sites. The Riparian site has the largest sample size of weights for male (`r juvenile_weights_sex_site_ref$sample_size[8]`), female (`r juvenile_weights_sex_site_ref$sample_size[7]`) and unknown (`r juvenile_weights_sex_site_ref$sample_size[9]`) juvenile hares, and the Black Spruce site has the fewest counts of weights for males (`r juvenile_weights_sex_site_ref$sample_size[2]`), females (`r juvenile_weights_sex_site_ref$sample_size[1]`) and unknown `r juvenile_weights_sex_site_ref$sample_size[3]`. The central tendency (mean) of female weights is below male weights at each site although there is a large spread (standard deviation), for Riparian females (`r juvenile_weights_sex_site_ref$mean[7]` ± `r juvenile_weights_sex_site_ref$sd[7]`; mean ± 1 standard deviation) compared to males (`r juvenile_weights_sex_site_ref$mean[8]`± `r juvenile_weights_sex_site_ref$sd[8]`). The biggest take away is that the large differences in sample size (n) and large standard deviations which impact the visible trends.


## Section 3: Juvenile weight comparison (male & female snowshoe hares)

```{r}
# Wrangle the data 
juvenile_m_f_weight <- bonanza_hares_data %>%
  filter(age == "j", 
         sex %in% c("m", "f")) %>% 
  select(sex, weight) %>% 
    mutate(sex = case_when(
    sex == "m" ~ "Male",
    sex == "f" ~ "Female"))
 
# Under the hood reference table

juvenile_m_f_weight_ref <- juvenile_m_f_weight %>% 
  group_by(sex) %>% 
  summarize(mean = mean(weight, na.rm = TRUE),
            median = median(weight, na.rm = TRUE),
            sd = sd(weight, na.rm = TRUE),
            sample_size = n())

# Finalized table 

juvenile_m_f_weight_summary <- juvenile_m_f_weight %>% 
  group_by(sex) %>% 
  summarize(mean = mean(weight, na.rm = TRUE),
            median = median(weight, na.rm = TRUE),
            sd = sd(weight, na.rm = TRUE),
            sample_size = n()) %>% 
  kable(caption = "Table 2: Descriptive statistics (mean, median, standard deviation and sample size) of juvenile weights (g) for male and femal snowshoe hares",
        col.names = c("Sex",
                      "Mean (g)",
                      "Median (g)",
                      "Standard deviation (g)",
                      "Sample size"),
        digits = 1) %>% 
  kable_minimal(html_font = "Georgia", font_size = 12, full_width = FALSE) %>% 
  kable_material(c("striped", "hover", "condensed"))
  
juvenile_m_f_weight_summary
```

```{r, include=FALSE}

# Exploratory graphs 

ggplot(data =juvenile_m_f_weight, aes(x = weight))+ # histogram
  geom_histogram()+
  facet_wrap(~sex)

ggplot(data =juvenile_m_f_weight, aes(sample = weight))+ # qq plot
  geom_qq()+
  facet_wrap(~sex)
 
# Sample size (363) in combination with histograms & QQ plots: I feel OK about using a t-test for a means comparison.         
```


```{r}

# Two sample t-test and Cohen test set up/ data wrangling

male_sample <- juvenile_m_f_weight %>% 
  filter(sex == "Male") %>% 
  drop_na(weight) %>% 
  pull(weight)


female_sample <- juvenile_m_f_weight %>% 
  filter(sex == "Female") %>% 
  drop_na(weight) %>% 
  pull(weight)
  


# Two sample t-test
m_f_weight_t <- t.test(male_sample,female_sample)

#m_f_weight_t # p = 0.7%

# Cohen test

m_f_weight_cohen <- cohen.d(male_sample, female_sample)

#m_f_weight_cohen
```

### Table 2 takeaways

On average male juvenile hares weigh more than females (`r juvenile_m_f_weight_ref$mean[2]` ± `juvenile_m_f_weight_ref$sd[2]` and `r juvenile_m_f_weight_ref$mean[1]`±`juvenile_m_f_weight_ref$sd[2]` g, respectively; mean ± 1 standard deviation). While the absolute difference in mean weight is `r round( juvenile_m_f_weight_ref$mean[2]-juvenile_m_f_weight_ref$mean[1], 2)`g ( a 10 % difference), the difference in means is significant (Welch's two-sample t-test: t(`r round(m_f_weight_t$parameter,2)`) = `r round(m_f_weight_t$statistic,2)`, p < 0.001), and the effect size is small (Cohen's *d* = `r round(m_f_weight_cohen$estimate,2)`) 

## Section 4: Relationship between juvenile weight & hind foot length

```{r fig.align='center'}
# Wrangle the data

juvenile_hfoot <- bonanza_hares_data %>% 
  filter(age == "j") %>% 
  select(weight, hindft)

# Data visualization

ggplot(data = juvenile_hfoot, aes(x = hindft, y = weight))+
  geom_jitter()+
  geom_smooth(method = "lm")+
  labs(x = "Hindfoot length (mm)",
       y = "Weight (g)",
       title = "Juvenile hare weight and hindfoot length (ZAUNER)")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title = element_text(face = "bold"))
```

```{r}
# Linear model stored as juvenile_lm 
juvenile_lm <- lm(weight ~ hindft, data = juvenile_hfoot)

# Return complere overview

#summary(juvenile_lm) 

# Use broom::tidy() function to see the model outputs in a nice format

juvenile_lm_tidy <- broom::tidy(juvenile_lm)

juvenile_lm_glance <- broom::glance(juvenile_lm)

# Find Pearson's r value which indicated the strength of the correlation between hind foot length and weight. 

juvenile_cor <- cor.test(juvenile_hfoot$weight, juvenile_hfoot$hindft)
```

***Figure 3:*** Relationship between juvenile hare weight (g) and hind food length (mm). Points indicate individual hare measurements. Linear model summary: $\beta$~1~ = `r round(juvenile_lm_tidy$estimate[2],2)` g/mm, p < 0.001,  R^2^ = `r round(juvenile_lm_glance$r.squared,2)`, Pearson's *r* = `r round(juvenile_cor$estimate,2)`. Based on the R^2^ value on 30% of variance in hare weights can be explained by hind foot length and Pearson's *r* value describes medium positive relationship between the two variables. Data: Kielland et al. (2017).

# Summary 

Exploratory data analysis reveals the following initial findings:

- The inconsistencies in sample sizes between years and sites make it difficult to draw definitive conclusions about the data.

- Given a small to moderate effect size and p value < 0.01, we can consider rejecting the null hypothesis that the means of 
- 

# Citations 

*Data citation*: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

