---
title: 'Task 2: Mini "exploratory report" on juvenile snowshoe hares in Bonanza Creek'
author: "Anna Zauner"
date: "11/4/2021"
output: 
  html_document: 
    theme: simplex
    number_sections: yes
---
# Introduction 

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.
 
*Data citation*: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

## Set up

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
```

## Read in the data

```{r}
bonanza_hares_data <- read_csv(here("data","bonanza_hares.csv")) %>% 
  clean_names() %>% 
  mutate(date = mdy(date))
```

# Data and anyalses

 A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.
 
# Juvenile hare initial investigation 

## Annual juvenile hare trap counts

. Count the total number of juvenile hare trappings during each year of the study, and create a finalized data visualization of the counts by year. Include a figure caption below the figure in your report. 

```{r}
annual_juvenile_trap_counts <- bonanza_hares_data %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(trap_counts = n())


ggplot(data = annual_juvenile_trap_counts, aes(x = year, y = trap_counts))+
  geom_bar(stat = "identity", fill = "blue")+
  labs(x = "Year",
       y = "Juvenile hare trap counts",
       title = "Annual juvenile hare trap counts (1998-2012)",
       caption = "Figure 1: Counts of juvenile hare trappings during each year of the study from 1998 to 2012")+
  theme_classic()+
  theme(plot.caption = element_text(hjust = 0.5, size = 8),
        plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(breaks = c(1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012))


```

```{r}





## Visualize juvenile hare weights

```{r}

## Juvenile weight comparison (male & female snowshoe hares)

```{r}

```

## Relationship between juvenile weight & hind foot length

```{r}

```

# Summary 

# Citations 

*Data citation*: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

