---
title: 'Task 2: Mini "exploratory report" on juvenile snowshoe hares in Bonanza Creek'
author: "Anna Zauner"
date: "11/4/2021"
output: 
  html_document: 
    theme: simplex
    number_sections: yes
---
# Introduction 

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.
 
*Data citation*: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

## Set up

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
```

## Read in the data

```{r}
bonanza_hares_data <- read_csv(here("data","bonanza_hares.csv")) %>% 
  clean_names() %>% 
  mutate(date = mdy(date))
```

# Data and anyalses

 A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.
 
# Juvenile hare initial investigation 

## Annual juvenile hare trap counts

. Count the total number of juvenile hare trappings during each year of the study, and create a finalized data visualization of the counts by year. Include a figure caption below the figure in your report. 

```{r}
annual_juvenile_trap_counts <- bonanza_hares_data %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(trap_counts = n())


ggplot(data = annual_juvenile_trap_counts, aes(x = year, y = trap_counts))+
  geom_bar(stat = "identity", fill = "blue")+
  labs(x = "Year",
       y = "Juvenile hare trap counts",
       title = "Annual juvenile hare trap counts (1998-2012)",
       caption = "Figure 1: Counts of juvenile hare trappings during each year of the study from 1998 to 2012")+
  theme_classic()+
  theme(plot.caption = element_text(hjust = 0.5, size = 8),
        plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(breaks = c(1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012))


```

### Figure 1 takeaways

There is no clear general trend displayed in figure 1, there is a large spread of annual juvenile hare trap counts, ranging from 38 to 410. The mean count of juvenile hares trapped annually is 213 and median is 179. The changes from year to year are stark and show no general trend, this could be due to the fact that these are absolute counts and are impacted by the effort of researchers. I would suggest standardizing the juvenile hare population by randomly selecting uniform sample sizes for each year. 

```{r, include=FALSE}
annual_counts_summary <- annual_juvenile_trap_counts %>% 
  summarize(mean = mean(trap_counts),
            median = median(trap_counts),
            minimum = min(trap_counts),
            maximum = max(trap_counts)) %>% 
  kable(caption = "Table 1: Summary statistics of annual juvenile hare trap counts",
        col.names = c("Mean",
                      "Median",
                      "Minimun",
                      "Maximum")) %>% 
  kable_minimal()
  
annual_counts_summary
```

## Visualize juvenile hare weights

```{r}
juvenile_weights_sex_site <- bonanza_hares_data %>% 
  group_by(grid)

ggplot(data = bonanza_hares_data, aes(x = weight))+
  geom_histogram()+
  aes(fill = sex)+
  facet_grid(grid ~ sex)+
  labs(title = "Juvenile hare weights by sex and site",
       x = "Weight (g)",
       y = "Count",
       caption = "Figure 2: Juvenile hare weights in grams recorded by sex (male, female or NA) and site recorded (Bonbs, Bonmat or Bonrip)",
       fill = "Sex")+
  theme(plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5),
        )
  


```
### Figure 2 takeaways


## Relationship between juvenile weight & hind foot length

```{r}

```

# Summary 

# Citations 

*Data citation*: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

